### 2023-01-27 definitions.tex
details

Since $\pred(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $I^{k+1}$ . By
Fact~\ref{fact basis}, we can compute a basis of $I^{k+1}$.
\end{proof}
\subsection{Defining resilience}

@@

\problemx{$k$-Resilience problem or Explicit Resilience problem}
{a transition system $(S,\rightarrow)$, an integer $k$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
{$\forall s' \in BAD, ~ s \rightarrow^* s' \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}

@@

\problemx{Bounded Resilience problem}
{a transition system $(S,\rightarrow)$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
{$\exists k \geq 0 ~ \forall s' \in BAD, ~ s \rightarrow^* s' \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}

\alain{repetition...} plus court: equivaut à demander si $\exists k \geq 0$ tel que $S$ est $k$-resilient.

\problemx{Bounded Resilience problem}
{a transition system $(S,\rightarrow)$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
{$\exists k \geq 0  S$ is $k$-resilient. ?\newline}

@@

\subsubsection{Resilience for WSTS}

@@

\problemx{$k$-resilience problem for WSTS}
{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $I$ with a given basis, a decidable downward-closed set $J$.}
{$\forall s' ~ (s \rightarrow^* s') \wedge s' \in J  \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}

@@

\problemx{bounded Resilience problem for WSTS}
{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}

>>>>

Since $\pred(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $I^{k+1}$ . By
Fact~\ref{fact basis}, we can compute a basis of $I^{k+1}$.
\end{proof}

%%%
\iffalse
%%%%
%%%%
\subsection{Defining resilience}

@@

\problemx{$k$-Resilience problem }
{a transition system $(S,\rightarrow)$, an integer $k$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
{$\forall s' \in BAD, ~ s \rightarrow^* s' \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}

@@

%
%\problemx{Bounded Resilience problem}
%{a transition system $(S,\rightarrow)$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
%{$\exists k \geq 0 ~ \forall s' \in BAD, ~ s \rightarrow^* s' \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}
%

\problemx{Bounded Resilience problem}
{a transition system $(S,\rightarrow)$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
{$\exists k \geq 0  S$ is $k$-resilient. ?\newline}
%%
\fi
%%%

@@

\section{Resilience for WSTS}

@@

\problemx{$k$-resilience problem for WSTS}
{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
{$\forall s' ~ (s \rightarrow^* s') \wedge s' \in D  \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
%
plus simple : la $(D,U)$-resilience: $D \rightarrow^{\leq k} U$ ?  \alain{ça signifie pour tout $d \in D$ il existe un chemin de longueur $\leq k$ qui arrive dans un $u \in U$. c'est plus simple et plus général et ça ne demande moins de décider l'accessibilité....comment faire pour vérifier pour tous les (D,U) ? sinon ça dépend tjs de l'état intial $s$ et s'il faut vérifier pour tous les $s$....}

@@

\problemx{bounded Resilience problem for WSTS}
{A state $s$ of a WSTS $W=(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
{$\exists k \geq 0$ such that $W$ is $k$-resilient. ?\newline}

### 2023-01-27 proof.tex
details

\section{Decidability of resilience for downward-WSTS}

@@

%		enlever l'hypothèse finitely branching ?
A downward-WSTS has the \emph{reflexive} compatibility if for all transition  $s_1 \rightarrow t_1$ with
$s_1, t_1 , s_2 \in S$ and $s_2 \leq s_1$, either $t_1 \leq s_2$ or there exists a transition $s_2 \rightarrow t_2$ with $t_2 \leq t_1$.

Downward-WSTS enjoy a nice reachability property.

@@

%%%%%
\alain{peut-on faire mieux ?}

\begin{lemma}\label{Lemma intersection}
Let $A \subseteq S$, $J \subseteq S$ downward-closed and $I \subseteq S$ upward-closed. 
Then $A \cap J \subseteq I$  iff $ (\uparrow  A) \cap J \subseteq I$.
\end{lemma}

If 
$A \cap J \subseteq I$,
then
${\uparrow (A \cap J)} \subseteq {\uparrow I} = I$.
However
$({\uparrow A}) \cap J \subseteq {\uparrow (A \cap J)}$.
Indeed if $x$ in $({\uparrow A} \cap J)$,
$x$ is in $\uparrow A$
so there exists $a \in A$ such that $x \geq a$.
However $x$ is also in $J$, meaning $a$ is in $J$.
Meaning $a$ is in $(A \cap J)$
and $x \in { \uparrow (A \cap J)}$.

In the other direction,
since $A \subseteq {\uparrow A}$,
$({\uparrow  A}) \cap J \subseteq I$ implies
$A \cap J \subseteq ({\uparrow  A}) \cap J \subseteq I$.
\end{proof}

\begin{corollary}
For all $k \in \N$,
$ post^*(s)\cap J \subseteq I^k $  iff $ (\uparrow  post^*(s)) \cap J \subseteq I^k$. 
% et travailler avec $\downarrow post^*(s)$ à la place de $\uparrow post^*(s)$.
\end{corollary}

@@

Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , . . . , i.e. upward closed subsets $D_i \subseteq S$ such that $\uparrow post(D_i ) \subseteq D_i$. 
Every inductive invariant $D_i$ is an “over-approximation” of $\uparrow post^*(s)$ if it contains $s$.
(on énumère des sur-approximations de la cloture par le haut de $post*(s)$ par leur bases finies).
Each “over-approximation” $D_i$ is given by its basis $b(D_i)$. Notice that, by standard monotonicity, $\uparrow post^*(s)$ is such an inductive invariant and may
eventually be found.
% \alain{attention D est clos par le bas et $\uparrow post^*(s)$ est clos par le haut, je ne vois pas pourquoi $\uparrow post^*(s) \subseteq D$}
% \mathieu{Oui effectivement les D devraient être clos par le haut, je corrige ça à la prochaine update.}

Procedure 1 stops when it finds a basis $b(D)$ of an invariant $D$ such that
$b(D)  \cap J \subseteq I^k$.  Since $b(D)$ is finite and $J$ is decidable, we can
directly compute $b(D)  \cap J$.
% We can compute a basis of $I^{k+1}$ if we have a basis of $I^k$.
We can compute a basis
of $I^{k+1}$ if we have a basis of $I^k$. %This follows by the proof of Lemma 3.
Due to the Lemma, 
$b(D)  \cap J \subseteq I^k$ implies
$D  \cap J \subseteq I^k$.
Hence
$b(D)  \cap J \subseteq I^k$ implies
$\uparrow post^*(s) \cap J \subseteq D  \cap J \subseteq I^k$.
(since $D$ contains $ \uparrow post^*(s)$).

@@

\begin{theorem}
{\sc $k$-resilience} is decidable for effective WSTS with downward/strong upward compatibility.
%	with strong monotony.
\end{theorem}

Assume  $(S, \rightarrow, \leq)$ is a WSTS with downward/upward compatibility, $J$ is a decidable downward-closed subset of $S$, and $I$ is an upward-closed set with a given basis.

We define inductively
$I^{k} = \bigcup_{0 \leq j \leq k} \pred^j(I)$. Note that for all $k \in \N$, $I^k$ is upward-closed due to
the strongly upward compatibility of $(S, \rightarrow, \leq)$.

In order to certify inclusion in $I^k$, we need to work with finite representations.
The next lemma uses that $I$ and $J$ are upward- and downward-closed, respectively.

@@

\begin{figure}
\fbox{\parbox[t][3.cm][c]{6cm}{
$\phantom{a}$\\
(1)\phantom{aaaaa} $i \leftarrow 0$\\
(2)\phantom{aaaaa}\textbf{while} $\neg( \uparrow post^*(D_i) \subseteq D_i $
			 and $ s \in D_i$
			 and $ b(D)  \cap J \subseteq I^k  )$ \textbf{loop} \newline
	(3)\phantom{aaaaaa}$\phantom{aaaa} i \leftarrow i +1$ \newline
(4)\phantom{aaaaa}\textbf{end loop} \newline
(5)\phantom{aaaaa}\textbf{return} $\text{\textit{false}}$ \newline
	}
}
	\caption{\textbf{Procedure 1:} enumerates inductive invariants to find an inclusion certificate.}\label{procedure1}
\end{figure}

\begin{figure}
\fbox{\parbox[t][3.cm][c]{6cm}{
$\phantom{a}$\\
(1)\phantom{aaaaa} $D \leftarrow \{ s \} $\\
(2)\phantom{aaaaa}\textbf{while} $D \cap J \subseteq I^k $ \textbf{loop} \newline
	(3)\phantom{aaaaaa}$\phantom{aaaa} D \leftarrow D \cup post(D)$ \newline
(4)\phantom{aaaaa}\textbf{end loop} \newline
(5)\phantom{aaaaa}\textbf{return} $\text{\textit{false}}$ \newline
	}
}
	\caption{\textbf{Procedure 2:} searches for a non-inclusion certificate.}\label{procedure2}
\end{figure}

\newpage

We show that these two procedures are correct:

\item $k$-resilience holds if, and only if, Procedure 1 terminates.

\item $k$-resilience do not hold if, and only if, Procedure 2 terminates.

\end{enumerate}

Proof:

@@

\begin{theorem}
{\sc bounded resilience} and {\sc resilience} are decidable for WSTS with downward/strong upward compatibility.
\end{theorem}

@@

Iteratively check whether $k$-resilience holds. 
If this is the case, return $k_{min} = k$. Otherwise check whether 
$I^{k+1} = I^k$. If so, return $-1$ (false), otherwise
continue.
The stop condition is decidable
and by Fact~\ref{stop condition} also sufficient. 
\end{proof}

{\sc resilience} is decidable for WSTS with strong upward compatibility and k-resilience.

jouer avec les hypothèses downward/upward compatibility

>>>>

\section{Resilience for downward-WSTS}

@@

%		enlever l'hypothèse finitely branching ?
A downward-WSTS has the \emph{reflexive} compatibility if for all transition  $s_1 \rightarrow t_1$ with
$s_1, t_1 , s_2 \in S$ and $s_2 \leq s_1$, either $t_1 \leq s_2$ or there exists a transition $s_2 \rightarrow t_2$ with $t_2 \leq t_1$.\\

Downward-WSTS with reflexive compatibility enjoy a nice reachability property.

@@

%%%%%
\alain{peut-on faire mieux ? enlever des hypothèses ?}

Let us establish a general Lemma on downward-closed and upward-closed sets.

\begin{lemma}\label{Lemma intersection}
Let $A \subseteq S$, $J \subseteq S$ be a downward-closed set and $I \subseteq S$ be an upward-closed set. 
Then $A \cap J \subseteq I$  iff $ (\uparrow  A) \cap J \subseteq I$.
\end{lemma}

\begin{proof}
Let us suppose that $A \cap J \subseteq I$. Then ${\uparrow (A \cap J)} \subseteq {\uparrow I} = I$.
Show that $({\uparrow A}) \cap J \subseteq {\uparrow (A \cap J)}$.
Let $x \in ({\uparrow A}) \cap J$, then there exists $a \in A$ such that $x \geq a$.
Since $x \in J$ and $J$ is downward-closed, we also have $a \in J$.
Hence $a \in A \cap J$ and then $x \in { \uparrow (A \cap J)}$.
In the other direction,
since $A \subseteq {\uparrow A}$, the inclusion
$({\uparrow  A}) \cap J \subseteq I$ implies
$A \cap J \subseteq ({\uparrow  A}) \cap J \subseteq I$.
\end{proof}

Let us apply the previous Lemma with $A=post^*(s)$.

\begin{corollary}
For all $k \in \N$ and $s \in S$,
$ post^*(s)\cap J \subseteq I^k $  iff $ (\uparrow  post^*(s)) \cap J \subseteq I^k$. 
% et travailler avec $\downarrow post^*(s)$ à la place de $\uparrow post^*(s)$.
\end{corollary}

@@

{\bf Explanations on the two procedures.}
Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , . . . , i.e. upward closed subsets $D_i \subseteq S$ such that $\uparrow post(D_i ) \subseteq D_i$. \alain{write U for upward closed subset and not D as for downward...}
Every inductive invariant $D_i$ is an “over-approximation” of $\uparrow post^*(s)$ if it contains $s$.
(on énumère des sur-approximations de la cloture par le haut de $post*(s)$ par leur bases finies).
Each “over-approximation” $D_i$ is given by its basis $b(D_i)$. Notice that, by standard monotonicity, $\uparrow post^*(s)$ is such an inductive invariant and may
eventually be found.
% \alain{attention D est clos par le bas et $\uparrow post^*(s)$ est clos par le haut, je ne vois pas pourquoi $\uparrow post^*(s) \subseteq D$}
% \mathieu{Oui effectivement les D devraient être clos par le haut, je corrige ça à la prochaine update.}

Procedure 1 stops when it finds a basis $b(D)$ of an invariant $D$ such that
$b(D)  \cap J \subseteq I^k$.  Since $b(D)$ is finite and $J$ is decidable, we can
directly compute $b(D)  \cap J$.
% We can compute a basis of $I^{k+1}$ if we have a basis of $I^k$.
We can compute a basis
of $I^{k+1}$ if we have a basis of $I^k$. %This follows by the proof of Lemma 3.
Due to the Lemma ????, 
$b(D)  \cap J \subseteq I^k$ implies
$D  \cap J \subseteq I^k$.
Hence
$b(D)  \cap J \subseteq I^k$ implies
$\uparrow post^*(s) \cap J \subseteq D  \cap J \subseteq I^k$.
(since $ \uparrow post^*(s)  \subseteq D$).

@@

@@

\begin{figure}
\fbox{\parbox[t][3.cm][c]{6cm}{
$\phantom{a}$\\
(1)\phantom{aaaaa} $i \leftarrow 0$\\
(2)\phantom{aaaaa}\textbf{while} $\neg( \uparrow post^*(D_i) \subseteq D_i $
			 and $ s \in D_i$
			 and $ b(D)  \cap J \subseteq I^k  )$ \textbf{loop} \newline
	(3)\phantom{aaaaaa}$\phantom{aaaa} i \leftarrow i +1$ \newline
(4)\phantom{aaaaa}\textbf{end loop} \newline
(5)\phantom{aaaaa}\textbf{return} $\text{\textit{false}}$ \newline
	}
}
	\caption{\textbf{Procedure RES:} enumerates inductive invariants to find an inclusion certificate.}\label{procedure1}
\end{figure}

\begin{figure}
\fbox{\parbox[t][3.cm][c]{6cm}{
$\phantom{a}$\\
(1)\phantom{aaaaa} $D \leftarrow \{ s \} $\\
(2)\phantom{aaaaa}\textbf{while} $D \cap J \subseteq I^k $ \textbf{loop} \newline
	(3)\phantom{aaaaaa}$\phantom{aaaa} D \leftarrow D \cup post(D)$ \newline
(4)\phantom{aaaaa}\textbf{end loop} \newline
(5)\phantom{aaaaa}\textbf{return} $\text{\textit{false}}$ \newline
	}
}
	\caption{\textbf{Procedure NOT-RES:} searches for a non-inclusion certificate.}\label{procedure2}
\end{figure}

We show that these two procedures allow to solve the $k$-resilience.

\begin{proposition}
The procedures $RES$ and $NOT-RES$ have the following property.
\begin{itemize}
%
\item $k$-resilience holds if and only if Procedure $RES$ terminates.
%
\item $k$-resilience do not hold if and only if Procedure $NOT-RES$ terminates.
%
\end{itemize}
\end{proposition}

@@

\begin{theorem}
{\sc $k$-resilience} is decidable for effective \alain{pas défini}WSTS with downward/strong upward compatibility.
%	with strong monotony.
\end{theorem}
\alain{les resultatts et preuves ne sont pas faciles à lire: faire un seul niveau de parentheses}
\begin{proof}
%			\begin{proof}

Assume  $(S, \rightarrow, \leq)$ is a WSTS with downward/upward compatibility, $J$ is a decidable downward-closed subset of $S$, and $I$ is an upward-closed set with a given basis.
We define inductively
$I^{k} = \bigcup_{0 \leq j \leq k} \pred^j(I)$. Note that for all $k \in \N$, $I^k$ is upward-closed due to
the strongly upward compatibility of $(S, \rightarrow, \leq)$.

%    In order to certify inclusion in $I^k$, we need to work with finite representations.
%    The next lemma \alain{where ? the previous lemmas ? numbers ?} 
%    uses that $I$ and $J$ are upward- and downward-closed, respectively.
\end{proof}

\begin{theorem}
{\sc bounded resilience} and {\sc resilience} are decidable for WSTS with downward/strong upward compatibility.
\alain{+effective hypothesis}
\end{theorem}

@@

Iteratively check whether $k$-resilience holds. 
If this is the case, return $k_{min} = k$. Otherwise check whether 
$I^{k+1} = I^k$. If so, return $-1$ (false), otherwise
continue.
The stop condition is decidable
and by Fact~\ref{stop condition} also sufficient. 
\end{proof}
%
%{\sc resilience} is decidable for WSTS with strong upward compatibility and k-resilience.
%
%jouer avec les hypothèses downward/upward compatibility
